Soc504 Replication Project Report

### 2015-05-12 replication_report-extensions.Rmd
final version of replication report

---
title: "Soc504 Replication Project Report"
author: "Yari & Kyonne" 
date: "May 10, 2015"
output: html_document
---

@@

```{r eval= FALSE}

@@

# Chi-Square, no correction, no simulation:
cs1 <- chisq.test(tbl, correct = FALSE)

@@

# Power analysis -- Find the effect size for a test with .8 sensitivity:
pwr.chisq.test(w = NULL, N = 91, df = 1, sig.level = 0.05, power = .8)

# Analyses on SUBSETTED sample:

# Filter out additional 19 Ss with too positive attitudes in "disliked music" condition (reported ratings above 3) & too negative attitudes in "liked music" condition (reported ratings below 3):

@@

@@

# Check those means (for comparison with article):
dat_e %>%
  group_by(music) %>%
  summarise(avg = mean(avg_att_muz))

# Exclude when average attitude > 3 in disliked cond. or average attitude < 3 for liked cond. (1 == liked music, 2 == disliked music)
dat_egood <- dat_e %>%
  filter(music == 1 & avg_att_muz > 3 | music == 2 & avg_att_muz < 3)

## Redo Chi-Square with subsetted data:
# Test whether music condition is independent of pen choice:
# Create 2x2 contingency table of frequencies - pen choice by music condition
tbl2 <- table(dat_egood$music, dat_egood$chose_advertised_pen) 
tbl2

# Chi-Square, no correction, no simulation:
cs2 <- chisq.test(tbl2, correct = FALSE)
tidy(cs2)

@@

# Confidence Intervals (not replicated from paper):
#upper (95%)
#CI95_up <- qchisq(.95, df = 1)
#lower (5%)
#CI95_low <- qchisq(.05, df = 1)

## Replicate Table: Frequency table & chi-square statistic of subsetted data
# (replicate "Experiment 3" portion of the table from article)

@@

### (3.1) Extension A: Graphing the frequency of responses in each condition.

@@

#### (3.1.1) Code for graph.

@@

#### (3.2.1) A quick check of N by gender in data.

@@

#Get counts and check values against article for correct coding scheme:
count(dat_egood, geslacht1)
#confirmed:
## 1 = male
## 2 = female

@@

#### (3.2.2) Code for logistical regressions.

#Recode dummy variable for music as 0 and 1 for easier interpretation of estimates:
Music <- factor(dat_egood$music,
                    levels = c(1,2),
                    labels = c("LikedMusic", "DislikedMusic"))

dat_egood$music <- as.numeric(dat_egood$music)
dat_egood <- dat_egood %>%
  mutate(music.recode2 = (music - 1))
View(dat_egood)

#Set factor names:
Music <- factor(dat_egood$music.recode2,
                    levels = c(0, 1),
                    labels = c("Liked", "DisLiked"))

ChosenPen <- factor(dat_egood$chose_advertised_pen,
                levels = c(1, 0),
                labels = c("ChoseAdPen", "NotChoseAdPen"))

#Compare counts of each gender in paper to data for coding:
count(dat_egood, geslacht1)
## 1 = male
## 2 = female

# need to check which is which:
Gender <- factor(dat_egood$geslacht1,
                    levels = c(1,2),
                    labels = c("Male", "Female"))

# for relationship between advertised pen and music, music 2 (disliked) seems significant:
fit1 <- glm(ChosenPen ~ Music, data = dat_egood, family = binomial)
fit1 <- tidy(fit1)
fit1

#### calculate odds 
fit1$estimate
fit1$estimate[1] + fit1$estimate[2] ## 1.216
## odds ratio
exp(fit1$estimate[2]) ##4.43
## predicted probabilities:
# Disliked Music:
(exp(fit1$estimate[1] + fit1$estimate[2])) / (1 + exp(fit1$estimate[1] + fit1$estimate[2]))
# Liked Music: 
exp(fit1$estimate[1]) / (1 + exp(fit1$estimate[1]))

# does the relationship change when looking ONLY at males & ONLY females?

male.dat_egood <- dat_egood %>%
  filter(geslacht1 == "1")
#count(male.dat_egood, geslacht1)

female.dat_egood <- dat_egood %>%
  filter(geslacht1 == "2")
#count(female.dat_egood, geslacht1)

# MALES:
fitmale <- glm(ChosenPen ~ Music, data = male.dat_egood, family=binomial)
fitmale <- tidy(fitmale)
fitmale
#### calculate odds 
fitmale$estimate
fitmale$estimate[1] + fitmale$estimate[2] ## 1.216
## odds ratio
exp(fitmale$estimate[2]) ##4.43

# FEMALES:
fitfemale <- glm(ChosenPen ~ Music, data = female.dat_egood, family=binomial)
fitfemale <- tidy(fitfemale)
fitfemale
#### calculate odds 
fitfemale$estimate
fitfemale$estimate[1] + fitfemale$estimate[2] ## 1.216
## odds ratio
exp(fitfemale$estimate[2]) ## 4.43

### predicted probabilities = confirmed, gender does not have a moderating effect on the likelihood 
# of choosing the advertised pen when listening to liked versus disliked music.
# Males - Disliked Music = 0.77
(exp(fitmale$estimate[1] + fitmale$estimate[2])) / 
  (1 + exp(fitmale$estimate[1] + fitmale$estimate[2])) 
# Males - Liked Music = 0.43
exp(fitmale$estimate[1]) / (1 + exp(fitmale$estimate[1]))

# Females - Disliked Music = 0.77
(exp(fitfemale$estimate[1] + fitfemale$estimate[2])) / 
  (1 + exp(fitfemale$estimate[1] + fitfemale$estimate[2])) 
# Males - Liked Music = 0.43
exp(fitfemale$estimate[1]) / (1 + exp(fitfemale$estimate[1]))

```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the figures. Further, the 'eval = FALSE' parameter was added to the top code chunk to prevent printing of the code along with any generated figures

>>>>

---
title: "Soc504 Replication Project Report"
author: "Yari & Kyonne" 
date: "May 12, 2015"
output: html_document
---

@@

```{r}

@@

# Chi-Square, no correction, no simulation:
cs1 <- chisq.test(tbl, correct = FALSE)
cs1 <- tidy(cs1)

@@

# Power analysis -- Find the effect size for a test with .8 sensitivity:
pwr.chisq.test(w = NULL, N = 91, df = 1, sig.level = 0.05, power = .8)
```

Chi-square analyses on the entire sample replicate the results reported in the article and confirm that participants in the liked and disliked music condition did not choose the advertised pen at an equal rate (chi-square = *`r cs1[1,1]`*, *p = `r cs1[1,2]`*).

```{r, eval=FALSE}

## Analyses on SUBSETTED sample:

# Filter out additional 19 Ss with too positive attitudes in "disliked music" condition (reported ratings above 3 in attitudes survey) & too negative attitudes in "liked music" condition (reported ratings below 3):

@@

# Check the grand/group-level means (for comparison with article):
dat_e %>%
  group_by(music) %>%
  summarise(avg = mean(avg_att_muz))

```{r}

# Exclude participants with average attitude > 3 in disliked condition or average attitude < 3 for liked condition (1 == liked music, 2 == disliked music)

@@

# Create a new dataframe "dat_egood" with filtered participants:
dat_egood <- dat_e %>%
  filter(music == 1 & avg_att_muz > 3 | music == 2 & avg_att_muz < 3)

## Redo Chi-Square with subsetted data:
# Test whether music condition is independent of pen choice:
# Create 2x2 contingency table of frequencies - pen choice by music condition
tbl2 <- table(dat_egood$music, dat_egood$chose_advertised_pen)

# Chi-Square, no correction, no simulation:
cs2 <- chisq.test(tbl2, correct = FALSE)
cs2 <- tidy(cs2)

@@

# Confidence Intervals (these were not reported in paper, just for curiosity):
#upper (95%)
#CI95_up <- qchisq(.95, df = 1)
#lower (5%)
#CI95_low <- qchisq(.05, df = 1)

Chi-square analyses on the subsetted sample replicate from the paper and further confirm that participants in the liked and disliked music condition did not choose the advertised pen at an equal rate (chi-square = *`r cs2[1,1]`*, *p = `r cs2[1,2]`*).

```{r, eval=FALSE}

## Replicate Table: Frequency table & chi-square statistic of subsetted data

# (replicate "Experiment 3" portion of the table from article)

@@

### (3.1) Extension A: Graphing the frequency of responses in each condition.
Note: the parameter "echo = FALSE" was added to this code chunk to prevent the printing of the code. For the code related to this graph, see section (3.1.1).

@@

#### (3.1.1) Code for graph.
Note: the "eval = FALSE" parameter was added to prevent the rendering of the figure. To observe the figure, see section (3.1).

@@

#### (3.2.1) A quick check of N by gender in data show a heavy proportion of females were sampled.

@@

#Get counts and check values against article for correct coding scheme:
count(dat_egood, geslacht1)
#confirmed coding scheme:
## 1 = male
## 2 = female

@@

#### (3.2.2) Code and analysis results for logistical regressions.

```{r}

# Recode dummy variable for music as 0 and 1 for easier interpretation of estimates:
dat_egood$music <- as.numeric(dat_egood$music)
dat_egood <- dat_egood %>%
  mutate(music.recode2 = (music - 1))

#Set factor names:
music_condition <- factor(dat_egood$music.recode2,
                    levels = c(0, 1),
                    labels = c("Liked", "DisLiked"))

pen_choice <- factor(dat_egood$chose_advertised_pen,
                levels = c(1, 0),
                labels = c("ChoseAdPen", "NotChoseAdPen"))

gender <- factor(dat_egood$geslacht1,
                    levels = c(1,2),
                    labels = c("Male", "Female"))

# for relationship between advertised pen and music:
fit1 <- glm(pen_choice ~ music_condition, data = dat_egood, family = binomial)
fit1 <- tidy(fit1)
fit1

#### calculate odds 
fit1$estimate[1] + fit1$estimate[2]

## odds ratio
exp(fit1$estimate[2])

## predicted probabilities:
# Disliked Music:
(exp(fit1$estimate[1] + fit1$estimate[2])) / (1 + exp(fit1$estimate[1] + fit1$estimate[2]))
# Liked Music: 
exp(fit1$estimate[1]) / (1 + exp(fit1$estimate[1]))

# Does the relationship change when looking ONLY at males & ONLY females?

male.dat_egood <- dat_egood %>%
  filter(geslacht1 == "1")
##double-check correct filtering:
#count(male.dat_egood, geslacht1)

female.dat_egood <- dat_egood %>%
  filter(geslacht1 == "2")
##double-check correct filtering:
#count(female.dat_egood, geslacht1)

# MALES:
fitmale <- glm(pen_choice ~ music_condition, data = male.dat_egood, family=binomial)
fitmale <- tidy(fitmale)
fitmale

#### calculate odds 
fitmale$estimate[1] + fitmale$estimate[2]
## odds ratio
exp(fitmale$estimate[2])

# FEMALES:
fitfemale <- glm(pen_choice ~ music_condition, data = female.dat_egood, family=binomial)
fitfemale <- tidy(fitfemale)
fitfemale

#### calculate odds 
fitfemale$estimate[1] + fitfemale$estimate[2]
## odds ratio
exp(fitfemale$estimate[2])

### predicted probabilities = confirmed, gender does not have a moderating effect on the likelihood 
# of choosing the advertised pen when listening to liked versus disliked music.
# Males - Disliked Music
(exp(fitmale$estimate[1] + fitmale$estimate[2])) / 
  (1 + exp(fitmale$estimate[1] + fitmale$estimate[2])) 
# Males - Liked Music
exp(fitmale$estimate[1]) / (1 + exp(fitmale$estimate[1]))

# Females - Disliked Music
(exp(fitfemale$estimate[1] + fitfemale$estimate[2])) / 
  (1 + exp(fitfemale$estimate[1] + fitfemale$estimate[2])) 
# Females - Liked Music
exp(fitfemale$estimate[1]) / (1 + exp(fitfemale$estimate[1]))

```
The predicted probabilities for participants' pen choice in either the liked or disliked conditions did not change after splitting the sample by gender. Males and females were equally likely to choose the same pen, regardless of their assigned condition. However, these results should be qualified by the observation that this data contained over triple the number of female participants as male participants.
